---
title: "Importing and pre-processing real estate data"
author: "Will Bidstrup"
date: "26/12/2018"
output: html_document
---

This is a working file to gather relevant real estate information from [[realestate.com.au](https://www.realestate.com.au/rent/).  

The objective is to create a data frame that can be used for exploratory analysis.  

Key variables to gather include;  

- Price per week 
- Number of bedrooms  
- Postcode
- NBN status


```{r, message = FALSE}
library(tidyverse) # For everything
library(rvest) # For some prettier themes
```

# Scraping from the web

The first step is to define search parameters on the website itself and run a search. In this example i am searching for 3 bedroom houses near Adelaide 5000 between 300 and 450 dollars per week.   

```{r}
url <- "https://www.realestate.com.au/rent/property-unitblock-villa-townhouse-unit+apartment-house-with-3-bedrooms-between-300-450-in-adelaide,+sa+5000/list-1"
```

```{r}
# Read in the webpage
webpage <- read_html(url)
```


# Applying the search across multiple pages  

Section TBD

# Collecting variables

By using the SelectorGadget to understand the html structure I select the variables of interest. Most then require a few additional pre-processing steps before they can be put together in a data frame.   

## Price per week

```{r}
# Pull from html
price_data_html <- html_nodes(webpage,'.priceText')

# Convert to text
price_data <- html_text(price_data_html)

# TODO - regex to strip to numeric only

```


## Number of bedrooms

Beds, bathrooms and car spaces are all in the same field. I only care about beds to start with.  

```{r}
# Pull from html
beds_data_html <- html_nodes(webpage,'dd')

# Convert to text
beds_data <- html_text(beds_data_html)

# Split into beds, baths, cars
beds <- as.integer(beds_data[seq(1, length(beds_data), 3)])
bathrooms <- as.integer(beds_data[seq(2, length(beds_data), 3)])
cars <- as.integer(beds_data[seq(3, length(beds_data), 3)]) # TODO Note - looks like one less - could be that this number does not exist in one of the listings?

```



## Postcode  

## NBN status


# Creating data frame
